Modern Graph Generators: A Survey

### 2018-09-12 generators_SocialNetworks.tex
Minor corrections

\paragraph{Realistic Social Network}
~\cite{Barrett:2009:GAL:1995456.1995598} focused on construction of
realistic social networks using a combination of public and private data sets
and large-scale agent based techniques. The process works as follows: Step 1
creates a synthetic population by integrating databases from commercial and
public sources. In Step 2, a set of activity templates are determined. Each
synthetic individual is assigned a 24-hour activity sequence including
geo-locations for each activity. To demonstrate the approach, the authors develop a synthetic population for the
United States that models every individual in the population. The synthetic
population is a set of geographically located people and households. Household
structure and demographics are derived from U.S. Census data. The activity
templates are  based on several thousand responses to an activity or time-use
survey. Demographic information for each person and location, a minute-by-minute
schedule of each person's activities, and the locations where these activities
take place is generated by a combination of simulation and data fusion
techniques. This information is captured by a dynamic social contact network. Similar methods for agent-based strategies have been reported in~\cite{Bernstein:2013:SAS:2499604.2499609}.

\paragraph{Linkage vs. Activity Graphs} ~\cite{Yao2011} distinguished between two
types of social network graphs -- the \emph{linkage graph}, where nodes stand
for the people in the social network and edges are their friendship links, and
the \emph{activity graph}, where nodes also stand for the people but edges stand
for their interactions. On the basis of analysis of
Flickr\footnote{\url{https://www.flickr.com/}} social links and
Epinions\footnote{\url{http://www.epinions.com/}} network of user interactions,
the authors discover that they both exhibit power-law degree distribution, high
clustering coefficient (community structure), and small diameter; also regarding
the dynamic properties they both follow densification law and relatively stable
clustering coefficient over time. However, the authors do not observe diameter
shrinking in Epinions? activity graph and there is a difference in degree
correlation (how frequently nodes with similar degrees connect to each other).
Namely linkage graphs have positive degree correlation whereas activity graphs
show neutral degree correlation. With regards to the findings, the proposed generator focusses on linkage graphs
with positive degree correlation. For this purpose it extends the ``forest
fire'' spreading process algorithm~\cite{Leskovec:2005:GOT:1081870.1081893} with
link symmetry. It has two parameters -- the \emph{burning probability} $P_b$
which is in charge of the burning process, and the \emph{symmetry probability}
$P_s$ which indicates backward linking from old nodes to new ones. $P_b$
controls a BFS-based forward burning process. The fire burns increasingly
fiercely with $P_b$ approaching 1. Meanwhile, $P_s$ adds fuel to the fire as it
brings more links. It gives chances for big nodes to connect back to big nodes.

@@

\paragraph{S3G2} The Scalable Structure-correlated Social Graph Generator
(S3G2)~\cite{Pham2013} is a general framework which  generates a directed
labeled graph, where the nodes are objects with property values, and their
structure is determined by the class a node belongs to. S3G2 does not aim at
generating near real-world data, but at generating synthetic graphs which a
correlated structure. It causes that the probability to choose a certain
property value (from a pre-defined dictionary), or the probability to connect
two nodes with an edge are influenced by existing data values. For example, it
is possible to have a correlated degree distribution, from which the degree of
each node is generated, correlated with properties of node. Hence the generator
can ensure that, e.g., people with many friends in a social network will
typically post more pictures than people with few friends, i.e., the amount of
friend nodes influences the amount of posted comment and picture nodes. The data generation process starts with generating a number of nodes with
property values generated according to specified property value correlations and
then adding respective edges according to specified correlation dimensions. It
has multiple phases, each focusing on one correlation dimension. Each pass along
one correlation dimension is a Map phase in which data is generated, followed by
a Reduce phase that sorts the data along the correlation dimension in the next
pass. A heuristic observation that ``the probability that two nodes are
connected is typically skewed with respect to some similarity between the
nodes'' enables to focus only on sliding window of most probable candidates. The core idea of the framework is demonstrated using an example of a social
network (consisting of persons and social activities).  The dictionaries for
property values are inspired by DBpedia and provided with 20 property value
correlations. The edges are generated according to 3 correlation dimensions.

@@

Attribute Synthetic Generator (ASG) is a network generator for reproducing the
node feature distribution of standard networks and rewiring the network to
preferentially connect nodes that exhibit a high feature similarity. The network
is initialized with a group of three nodes, and new nodes and links are added to
the network based on link density, homophily, and feature similarity. As new
nodes are created, their labels are assigned based on the prior label
distribution. After the network has reached the same number of nodes as the
original social media dataset, each node initially receives a random attribute
assignment. Then a stochastic optimization process is used to move the initial
assignments closer to the target distribution extracted from social media
dataset using the Particle Swarm Optimization algorithm. The tuned attributes
are then used to add additional links to the network based on the feature
similarity parameter -- a source node is selected randomly and connected to the
most similar node. Multi-Link Generator (MLG) further  uses link co-occurrence statistics from the
original dataset to create a multiplex network. MLG uses the same network growth
process as ASG.. Based on the link density parameter, either a new node is
generated with a label based on the label distribution of the target dataset or
a new link is created between two existing nodes.

\paragraph{LDBC SNB} The Social Network Benchmark
(SNB)~\cite{Erling:2015:LSN:2723372.2742786} provided by LDBC consists of three
distinct benchmarks on a common dataset corresponding to three different
workloads. SNB models a social network akin to Facebook. The dataset consists of
persons and a friendship network that connects them; whereas the majority of the
data is in the messages that these persons post in discussion trees on their
forums. The three query workloads involve: (1) SNB-Interactive (i.e., complex
read-only queries, that touch a significant amount of data), (2) SNB-BI (which
accesses a large percentage of all entities in the dataset and groups these in
various dimensions), and (3) SNB-Algorithms (i.e., graph analysis algorithms,
including PageRank, Community Detection, Clustering and Breadth First Search).
The graph generator realizes power laws, uses skewed value distributions, and
introduces plausible correlations between property values and graph structures.
It is implemented on top of Hadoop to provide scalability.

@@

\paragraph{Towards More Realistic Data} \cite{Nettleton2016} argued that the main body of existing work lies in
topology generation which approximate the characteristics of a real social
network (such as a small graph diameter, small average path length, skew degree
distribution, and community structures), however without data. Hence, they
introduced a general stochastic modeling system which allows the users to
populate a graph topology with data. The approach has three steps: (1) topology
generation (using R-MAT) plus community identification using the Louvain
method~\cite{1742-5468-2008-10-P10008} or usage of a real-world topology from
SNAP\footnote{\url{https://snap.stanford.edu/data/}}, (2) data definition
(following distribution profiles, attribute value definitions, using a
parameterizable set of data propagation rules and affinities), and (3) data
population.

>>>>

\paragraph{Realistic Social Network}
~\cite{Barrett:2009:GAL:1995456.1995598} focused on construction of
realistic social networks using a combination of public and private data sets
and large-scale agent based techniques. The process works as follows: In the first step
it creates a synthetic population by integrating databases from commercial and
public sources. In the second step, a set of activity templates are determined. Each
synthetic individual is assigned a 24-hour activity sequence including
geolocations for each activity. To demonstrate the approach, the authors develop a synthetic population for the
United States that models every individual in the population. The synthetic
population is a set of geographically located people and households. Household
structure and demographics are derived from U.S. Census data. The activity
templates are  based on several thousand responses to an activity or time-use
survey. Demographic information for each person and location, a minute-by-minute
schedule of each person's activities, and the locations where these activities
take place is generated by a combination of simulation and data fusion
techniques. This information is captured by a dynamic social contact network. Similar methods for agent-based strategies have been reported in~\cite{Bernstein:2013:SAS:2499604.2499609}.

\paragraph{Linkage vs. Activity Graphs} \cite{Yao2011} distinguished between two
types of social network graphs -- the \emph{linkage graph}, where nodes stand
for the people in the social network and edges are their friendship links, and
the \emph{activity graph}, where nodes also stand for the people but edges stand
for their interactions. On the basis of analysis of
Flickr\footnote{\url{https://www.flickr.com/}} social links and
Epinions\footnote{\url{http://www.epinions.com/}} network of user interactions,
the authors discover that they both exhibit power-law degree distribution, high
clustering coefficient (community structure), and small diameter; also regarding
the dynamic properties they both follow densification law and relatively stable
clustering coefficient over time. However, the authors do not observe diameter
shrinking in Epinions activity graph and there is a difference in degree
correlation (how frequently nodes with similar degrees connect to each other).
Namely linkage graphs have positive degree correlation whereas activity graphs
show neutral degree correlation. With regards to the findings, the proposed generator focusses on linkage graphs
with positive degree correlation. For this purpose it extends the forest
fire spreading process algorithm~\cite{Leskovec:2005:GOT:1081870.1081893} with
link symmetry. It has two parameters -- the \emph{burning probability} $P_b$
which is in charge of the burning process, and the \emph{symmetry probability}
$P_s$ which indicates backward linking from old nodes to new ones. $P_b$
controls a BFS-based forward burning process. The fire burns increasingly
fiercely with $P_b$ approaching 1. Meanwhile, $P_s$ adds fuel to the fire as it
brings more links. It gives chances for big nodes to connect back to big nodes.

@@

\paragraph{S3G2} The Scalable Structure-correlated Social Graph Generator
(S3G2)~\cite{Pham2013} is a general framework which  generates a directed
labeled graph, where the nodes are objects with property values, and their
structure is determined by the class a node belongs to. S3G2 does not aim at
generating near real-world data, but at generating synthetic graphs with a
correlated structure. It causes that the probability to choose a certain
property value (from a pre-defined dictionary), or the probability to connect
two nodes with an edge are influenced by existing data values. For example, it
is possible to have a correlated degree distribution, from which the degree of
each node is generated, correlated with properties of node. Hence the generator
can ensure that, e.g., people with many friends in a social network will
typically post more pictures than people with few friends, i.e., the amount of
friend nodes influences the amount of posted comment and picture nodes. The data generation process starts with generating a number of nodes with
property values generated according to specified property value correlations and
then adding respective edges according to specified correlation dimensions. It
has multiple phases, each focusing on one correlation dimension. Each pass along
one correlation dimension is a Map phase in which data is generated, followed by
a Reduce phase that sorts the data along the correlation dimension in the next
pass. A heuristic observation that ``the probability that two nodes are
connected is typically skewed with respect to some similarity between the
nodes'' enables to focus only on sliding window of most probable candidates. The core idea of the framework is demonstrated using an example of a social
network (consisting of persons and social activities).  The dictionaries for
property values are inspired by DBpedia and provided with 20 property value
correlations. The edges are generated according to 3 correlation dimensions.

@@

Attribute Synthetic Generator (ASG) is a network generator for reproducing the
node feature distribution of standard networks and rewiring the network to
preferentially connect nodes that exhibit a high feature similarity. The network
is initialized with a group of three nodes, and new nodes and links are added to
the network based on link density, homophily, and feature similarity. As new
nodes are created, their labels are assigned based on the prior label
distribution. After the network has reached the same number of nodes as the
original social media dataset, each node initially receives a random attribute
assignment. Then a stochastic optimization process is used to move the initial
assignments closer to the target distribution extracted from social media
dataset using the Particle Swarm Optimization algorithm. The tuned attributes
are then used to add additional links to the network based on the feature
similarity parameter -- a source node is selected randomly and connected to the
most similar node. Multi-Link Generator (MLG) further  uses link co-occurrence statistics from the
original dataset to create a multiplex network. MLG uses the same network growth
process as ASG. Based on the link density parameter, either a new node is
generated with a label based on the label distribution of the target dataset or
a new link is created between two existing nodes.

\paragraph{LDBC SNB} The Social Network Benchmark
(SNB)~\cite{Erling:2015:LSN:2723372.2742786} provided by LDBC consists of three
distinct benchmarks on a common dataset corresponding to three different
workloads. SNB models a social network akin to Facebook. The dataset consists of
persons and a friendship network that connects them; whereas the majority of the
data is in the messages that these persons post in discussion trees on their
forums. The three query workloads involve: (1) SNB-Interactive, i.e., complex
read-only queries, that touch a significant amount of data, (2) SNB-BI which
accesses a large percentage of all entities in the dataset and groups these in
various dimensions, and (3) SNB-Algorithms, i.e., graph analysis algorithms,
including PageRank, Community Detection, Clustering and Breadth First Search.
The graph generator realizes power laws, uses skewed value distributions, and
introduces plausible correlations between property values and graph structures.
It is implemented on top of Hadoop to provide scalability.

@@

\paragraph{Towards More Realistic Data} \cite{Nettleton2016} argued that the main body of existing work lies in
topology generation which approximates the characteristics of a real social
network (such as a small graph diameter, small average path length, skew degree
distribution, and community structures), however without data. Hence, they
introduced a general stochastic modeling system which allows the users to
populate a graph topology with data. The approach has three steps: (1) topology
generation (using R-MAT) plus community identification using the Louvain
method~\cite{1742-5468-2008-10-P10008} or usage of a real-world topology from
SNAP\footnote{\url{https://snap.stanford.edu/data/}}, (2) data definition
following distribution profiles, attribute value definitions, using a
parameterizable set of data propagation rules and affinities, and (3) data
population.

### 2018-09-12 generators_analytics.tex
Minor corrections

\paragraph{HPC Scalable Graph Analysis Benchmark} The HPC Scalable Graph
Analysis Benchmark~\cite{HPCgraph,Bader:2005:DIH:2099301.2099360} represents an
application with multiple analysis techniques that access a single data
structure representing a weighted, directed graph. The benchmark is composed of
four separated operations (graph construction, classification of large vertex
sets, graph extraction with BFS, and graph analysis with betweenness centrality)
on a graph that follows a power-law distribution. The graph generator constructs a list of edge tuples containing vertex
identifiers (with the edge direction from the first one to the second one) and
weights that represent data assigned to the edges of the multigraph in the form
of positive integers with a uniform random distribution. The generator has the
following parameters: number of vertices, number of edges, and maximum weight of
an edge. The algorithm of the generator is based on R-MAT~\cite{DBLP:conf/sdm/ChakrabartiZF04}. Since the authors aim
to avoid data locality, in the final step the vertex numbers are randomly
permuted, and then edge tuples randomly shuffled. A related project from the same authors developed for the 9th DIMACS Shortest
Paths Challenge is GTgraph~\cite{GTgraph}. It involves three types of graphs:
input graph instances used in the DARPA HPCS SSCA\#2 graph theory benchmark
(version 1.0), Erd\"{o}s-R\'{e}nyi random graphs, and small-world graphs based
on the R-MAT model~\cite{DBLP:conf/sdm/ChakrabartiZF04}.

>>>>

\TODO{A short intro.}

\paragraph{HPC Scalable Graph Analysis Benchmark} The HPC Scalable Graph
Analysis Benchmark~\cite{HPCgraph,Bader:2005:DIH:2099301.2099360} represents an
application with multiple analysis techniques that access a single data
structure representing a weighted, directed graph. The benchmark is composed of
four separated operations (graph construction, classification of large vertex
sets, graph extraction with BFS, and graph analysis with betweenness centrality)
on a graph that follows a power-law distribution. The graph generator constructs a list of edge tuples containing vertex
identifiers (with the edge direction from the first one to the second one) and
weights that represent data assigned to the edges of the multigraph in the form
of positive integers with a uniform random distribution. The generator has the
following parameters: number of vertices, number of edges, and maximum weight of
an edge. The algorithm of the generator is based on R-MAT. Since the authors aim
to avoid data locality, in the final step the vertex numbers are randomly
permuted, and then edge tuples randomly shuffled. A related project from the same authors developed for the 9th DIMACS Shortest
Paths Challenge is GTgraph~\cite{GTgraph}. It involves three types of graphs:
input graph instances used in the DARPA HPCS SSCA\#2 graph theory benchmark
(version 1.0), Erd\"{o}s-R\'{e}nyi random graphs, and small-world graphs based
on the R-MAT model.

\paragraph{Graphalytics} Graphalytics\footnote{\url{http://graphalytics.ewi.tudelft.nl/}} is an industrial-grade benchmark for graph analysis platforms~\cite{Iosup:2016:LGB:3007263.3007270}. It involves 6 real-world datasets and 2 synthetic datasets generated so that they cover two commonly used types of graphs: social network graphs generated using LDBC SNB graph generator (see Section~\ref{sec:generators_socialnetworks}) and power-law graphs generated by Graph500 (see Section~\ref{sec:generators_general}). The benchmark workload consists of 6 deterministic algorithms: breadth-first search, PageRank, weakly connected components, community detection using label propagation, local clustering coefficient, and single-source shortest paths. The benchmark uses various metrics to measure the performance and throughput for systems under test such as \emph{upload time} that measures the required time to preprocess and convert the graph into a suitable format for a graph processing system, or \emph{Makespan} that measures the total  execution time of a benchmarking workload algorithm. The benchmark also describes experiments for measuring the scalability of the systems under test. To facilitate the end user job of running the designed experimental workload, the benchmark provides  a performance evaluation framework, Granula\footnote{\url{https://github.com/atlarge-research/granula}}.%, that consists of three main components: modeler, archiver, and the visualizer.

### 2018-09-12 generators_community_detection.tex
Refactored graph streaming and community detection into others section
graph stream generator
Minor corrections

\subsection{Graphs Generators for Benchmarking Community Detection Algorithms}
\label{sec:generators_community_detection}

Community detection is one of the many graph analytics algorithms typically used on domains such as social networks or bioinformatics. Communities are groups of vertices that are highly connected among them, while being scarcely connected to the rest of the graph. Such communities emerge from the fact that real graphs are not random, but follow realworld dynamics that make similar entities to have a larger probability to be connected. As a consequence, detected communities are used used to reval vertices with similar characteristics, for instance to discover functionally equivalent proteins in protein-protein interaction networks, or persons with similar interests in social networks.

Such applications have made community detection a hot topic during the last fifteen years with tens of developed algorithms and detection strategies. In order to compare the quality of the different proposed techniques, one needs graphs with reference communities, that is,. communities which known beforehand. Since it is very difficult to have large real graphs with reference communities --mainly because these would require a manual labeling--, graphs for benchmarking community detection algorithms typically generated synthetically.

The first attempts to compare community detection algorithms using synthetic graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi~\cite{Erdos:1960} subgraphs, connected more internally than externally~\cite{danon2005comparing}. Each of these subgraphs has the same size and the same internal/external density of edges. However, such graphs miss the realism observed in real graphs, where communities are of different sizes and densities.

Thus, Lancichinetti, Fortunato and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark graphs for community detection where communities are of different sizes and densities. The generator generates communities of different sizes following a power-law distribution whose parameters can be configured. The degree of the nodes is also sampled from a power-law distribution. Additionally, the generator introduces the concept of the ``mixing factor'', which specifies the fraction of its links connecting nodes belonging to different communities. Such parameter allows to tune the degree of modularity of the generated graph, thus testing the robustness of hte algorithms under different conditions. The generation process starts with an empty graph and incrementally fills in the adjacency matrix by obeying the described constraints.

Lancichinetti, Fortunato and Radicchi \cite{PhysRevE.80.016118} extended LFR to support the notion of directed graphs and overlapping communities. Overlapping communities extend the notion of communities by allowing the sharing of vertices, thus a vertex can belong to more than one community. This extended generator allows controlling the same parameters of the previous version, but also the amount of overlap of the generated communities.

Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining} proposed hte use of real graphs with explicit group annotations (e.g. forums in a social network, categories of products, etc.) to infer what they call ``meta-communities'', and use them to evaluate overlapping community detection algorithms. However, a recent study from Hric, Darst and Fortunato~\cite{hric2014community} reveal a loose correspondence between  communities --the authors refer to them as structural communities-- and meta-communities.  This result reveals that those algorithms working for structural communities do not work well for finding meta-communities and viceversa, suggesting significantly different underlying characteristics between the two types of communities, which are yet to be indentified.

>>>>

\subsubsection{Graphs Generators for Benchmarking Community Detection Algorithms}
\label{sec:generators_community_detection}

Community detection is one of the many graph analytics algorithms typically used on domains such as social networks or bioinformatics. Communities are groups of vertices that are highly connected among them, while being scarcely connected to the rest of the graph. Such communities emerge from the fact that real graphs are not random, but follow real-world dynamics that make similar entities to have a larger probability to be connected. As a consequence, detected communities are used to reveal vertices with similar characteristics, for instance to discover functionally equivalent proteins in protein-protein interaction networks, or persons with similar interests in social networks. Such applications have made community detection a hot topic during the last fifteen years with tens of developed algorithms and detection strategies. In order to compare the quality of the different proposed techniques, one needs graphs with reference communities, that is, communities known beforehand. Since it is very difficult to have large real graphs with reference communities (mainly because these would require a manual labeling), graphs for benchmarking community detection algorithms are typically generated synthetically.

The first attempts to compare community detection algorithms using synthetic graphs proposed the use of random graphs composed by several Erd\"{o}s-R\'{e}nyi subgraphs, connected more internally than externally~\cite{danon2005comparing}. Each of these subgraphs has the same size and the same internal/external density of edges. However, such graphs miss the realism observed in real graphs, where communities are of different sizes and densities. Thus, Lancichinetti, Fortunato and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark graphs for community detection where communities are of different sizes and densities. The generator generates communities of different sizes following a power-law distribution whose parameters can be configured. The degree of the nodes is also sampled from a power-law distribution. Additionally, the generator introduces the concept of the ``mixing factor'', which specifies the fraction of its links connecting nodes belonging to different communities. Such parameter allows to tune the degree of modularity of the generated graph, thus testing the robustness of the algorithms under different conditions. The generation process starts with an empty graph and incrementally fills in the adjacency matrix by obeying the described constraints. Lancichinetti, Fortunato and Radicchi \cite{PhysRevE.80.016118} extended LFR to support the notion of directed graphs and overlapping communities. Overlapping communities extend the notion of communities by allowing the sharing of vertices, thus a vertex can belong to more than one community. This extended generator allows controlling the same parameters of the previous version, but also the amount of overlap of the generated communities.

Besides synthetic graph generators, Yang and Leskovec~\cite{yang2015defining} proposed hte use of real graphs with explicit group annotations (e.g. forums in a social network, categories of products, etc.) to infer what they call ``meta-communities'', and use them to evaluate overlapping community detection algorithms. However, a recent study from Hric, Darst and Fortunato~\cite{hric2014community} reveal a loose correspondence between  communities --the authors refer to them as structural communities-- and meta-communities.  This result reveals that those algorithms working for structural communities do not work well for finding meta-communities and viceversa, suggesting significantly different underlying characteristics between the two types of communities, which are yet to be identified.
